---
output: html_document
editor_options: 
  chunk_output_type: console
---

### Hazem Sayed
### 20/07/2018
Exploratory Data Analysis Nanodegree Project
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# 
library(ggplot2)
library(reshape2)
library(GGally)
library(plyr)
library("tidyverse")
library("gridExtra")
library("memisc")
library(dplyr)
library(corrplot)


```

```{r echo=FALSE, Load_the_Data, message=FALSE, warning=FALSE}
# Load the Data
df <- read.csv('../AppleStore.csv')
```

# Introduction about the Data set!
In this document we are going to explore the top trending apps in iOS app store.
The dataset has more than 7000 Apple iOS mobile application details. 
The data was extracted from the iTunes Search API at the Apple Inc website.

# Univariate Plots Section

```{r echo=FALSE,show.text=TRUE, Univariate_Plots, message=FALSE, warning=FALSE}
# Showing summary about the data structures
dim (df)
str(df)

# Adding a new variable to include size by MB.
df$size_mb <- df$size_bytes/1024/1024

# Showing frequency of apps in our data set sorted by their count.
qplot(x=fct_infreq(df$prime_genre), data = df) +
  labs (title="Frequncy of Apps based on the category",
        x="Categories",y="Count") +
  theme(axis.text.x=element_text(angle=90,hjust=1))
```

From the above histogram, we see that the top app category in our data is Games.
We can investigate better by deviding categories; Games, General App')

Let's check a quick summary about free and paid apps

```{r echo=FALSE,show.text=TRUE, Univariate_Plots2, message=FALSE, warning=FALSE}

# Creating a new variable is_game to seperate General apps and Games.
df$is_game <- df$prime_genre == 'Games'
df$is_game <- ifelse(df$is_game == TRUE, "Game", "General App")

# Showing summary about the price, also adding a new variable to seperate paid
# and free apps.
summary(df$price==0)
df$is_free <- df$price == 0
df$is_free <- ifelse(df$is_free == TRUE, "Free", "Paid")

```

Will be a good option to add a new feature to check if it's paid or a free app.

* About 60% of our apps data are recorded for free applications, the rest is 
paid.

```{r echo=FALSE,show.text=TRUE, Univariate_Plots3, message=FALSE, warning=FALSE}
# Showing frequency of apps in our data set based on app content category.
qplot(x=cont_rating, data = df) +
  labs (title="Frequncy of Apps based on the content type",
        x="Content Type",y="Count") 
```
 
Most of our data focus on applications that are for +4 ages. 
That makes a good  sense since we already have most of the data category 
are games which are +4 age

```{r echo=FALSE,show.text=TRUE, Univariate_Plots4, message=FALSE, warning=FALSE}
# Showing frequncy of Apps based on the user rating above 0
qplot(x=user_rating, data = subset(df,df$user_rating>0),binwidth = 1) + 
labs (title="Frequncy of Apps based on the user rating above 0",
      x="User Rating",y="Count") 

#df <- subset(df,df$user_rating>0)
```

We excluded 0 rating from our apps to better investigate the dataset. 
Most of apps are rated 4.5 and very few are rated as average 5.

```{r echo=FALSE,show.text=TRUE, Univariate_Plots5, message=FALSE, warning=FALSE}

# Create a new variable in our data set that include copy of the data 
# after seperating Games and non-Games.
dfGames <-subset(df,df$is_game == 'Game')
dfGeneral <-subset(df,df$is_game != 'Game')

# Showing top 10 Games based on user rating and total number of ratings
print("Top 10 Games based on user rating and total number of ratings")
head(dfGames[order(dfGames$user_rating,dfGames$rating_count_tot, decreasing= T),
c("track_name","rating_count_tot")], n = 10,data = dfGames)

# Showing top 10 General apps based on user rating and total number of ratings
print("Top 10 General apps based on user rating and total number of ratings")
head(dfGeneral[order(dfGeneral$user_rating,dfGeneral$rating_count_tot, 
                     decreasing= T),c("track_name","rating_count_tot",
                                      "prime_genre")], n = 10,data = dfGeneral)
# This results are sorted based on user rating first and total ratings count.

# Showing top 10 apps based on total count of user rating
print("Top 10 apps based on total count of user rating")
head(df[order(df$rating_count_tot, decreasing= T),
        c("track_name","rating_count_tot")], n = 10,data = df)
# This results are sorted based on user rating first and total ratings count.
```

# Univariate Analysis

### Structure of the dataset:

There are 7197 observation about apps in the US iOS app store in our dataset 
with 16 features.
Categorical Variables Are: prime_genre, currency, user_rating, cont_rating, 
is_game.
Numerical Variables Are: price, rating_count_tot, size_bytes, size_mb, 
user_rating_ver, ver, rating_count_ver.

The primary category for most of the observation are about Games, the rest are 
for general apps.

Price can be devided into Free or Paid.

75% of apps are less than $2 in price and have average rating of 4.5
- 

### Main feature(s) of interest in the dataset?

The main features of interest in our dataset are the app details like 
price, and user rating.

### Other features in the dataset I think will help support my \
investigation into the feature(s) of interest?

Lanugage Supported and size may be useful in future state to determine the 
relation between ratings.

All currencies in our dataset are USD. Which means no need to check any other 
currency for this oarticular dataset.

### What about creating new variables in the dataset?

I created app_size by mb to better understand the size of the app and 
how it relects to the user ratings. 
I also created a variable of "Is free?" to better understand if the app is paid
or free will reflect into the user ratings?
I created a new variable is_game .. to cateorize application 
based if it's a general app or a game.

I subset the data to exclude data for 0 number ratings. It's not fair to 
calculate zero ratings in our investigations.

> **GGally**: Let's have an overview of all variables together, that should 
give us a quick look to the relation between each variable and the rest of them.

```{r echo=FALSE, ggpairs, message=FALSE, warning=FALSE,fig.width=12, fig.height=12}
ggpairs(df[,c('price','rating_count_tot',
              'rating_count_ver','user_rating','user_rating_ver',
              'size_mb','is_game','is_free',
              'lang.num','sup_devices.num')]) + theme(axis.text.x=element_text(angle=90,hjust=1))
```

> **ggcorrplot**: A quick look to correlation between our numeric columns.

```{r echo=FALSE, ggcorrplot, message=FALSE, warning=FALSE}
M <-cor(select_if(df[,5:20], is.numeric))
corrplot(M, method="circle")
```

# Bivariate Plots Section

Is there a relation between the price of an app and it's mean user rating? 
Let's deep onto our data set to answer this question ^^

```{r echo=FALSE, Bivariate_Plots, message=FALSE, warning=FALSE,fig.width=8, fig.height=8}
# A plotting point showing the relation between price of an app 
# and it's mean user rating.
ggplot(aes(x=user_rating,y=price), data = subset(df,price<50)) +
geom_point() + xlim(c(1,5)) +
labs (x="Average User Rating", y="Price",
title="Not strong relation between price of an app and it's mean user rating.")+
facet_wrap(~is_game) 
```

In the above plot, I included prices less than 50 to exclude some few 
applications their prices over $100. As we see there are some outer points 
in general apps for 4-5 ratings, which fairly indicates that 
user rating/quality for a general app may be the reason for increasing the price. 
This rarely could be happen (4% acuurancy).

```{r echo=FALSE, Bivariate_Plots2, message=FALSE, warning=FALSE}
# A plot line showing the relation between number of languages supported 
        #by an app and it's quality/user rating.
ggplot(aes(x=user_rating,y=lang.num), data = df) +
  geom_line(stat = 'summary', fun.y = mean) +
  labs (x="Average User Rating", y="Number of language supported",
        title="The relation between number of languages supported 
        by an app and it's quality/user rating")

summary(df$lang.num)
```

Yes, number of lanuages supported by an app may affect on it's total rating. 
We will check whether it affects on the price or not in a further analysis.

75% of our data set have 8 language supported or less. 
Most of them have only one language supported.

```{r echo=FALSE, Bivariate_Plots10, message=FALSE, warning=FALSE}
# Showing the relation between number of devices supported by an app and 
        #it's quality/user rating
ggplot(aes(y=user_rating,x=df$sup_devices.num), data = df) +
  scale_x_continuous(limits=c(36,47),breaks = seq(36,47,1)) +
  geom_line(stat = 'summary', fun.y = mean) + 
  labs (y="Average User Rating", x="Number of devices supported",
        title="The relation between number of devices supported by an app and 
        it's quality/user rating")

# Printing summary and coorelation between user rating and devices supported.
summary(df$sup_devices.num)
cor(df$user_rating,df$sup_devices.num)
```

The minumum number of supported devices is 9. Median is 37 

75% of our data set have 39 language supported or less. Most of them have 37 
devices supported.

Correlation is negative 4/100.

```{r echo=FALSE, Bivariate_Plots3, message=FALSE, warning=FALSE}
# a boxlot plot for content rating as a categorical value and price for an app.
ggplot(aes(x=cont_rating,y=price), data =subset(df,df$price>0)) +
  ylim(0,25) +
  geom_boxplot() 

summary.factor(df$cont_rating,stats=TRUE)
```

All content based apps are nearly have 3 in user rating. 
There are some slight diffferences but not that much to decide that the content 
rating may affect on the user ratings! 

* 25% of 12+ apps have prices less than $5, 
* 9+ and 17+ are nearly the same in regarding of the price.
* 4+ apps which include high number of games are less than $3.
There are some ouliers for apps high than $7.


Q: Which category has the most highgest rating?

```{r echo=FALSE, Bivariate_Plots4, message=FALSE, warning=FALSE,fig.width=12, fig.height=12}
# A histogram showing the relation between user ratings and an app category.
ggplot(aes(x=prime_genre,y=user_rating,fill=prime_genre), data = df) +
  geom_histogram(stat = 'summary', fun.y = mean) + coord_flip() +
  labs (x="Categories", y="Average of user rating",
        title="Does mean user ratings  depend on app category?") + 
  facet_wrap(~is_free) + theme_minimal()

```

* Productivity and Music have the highest average rating in free apps.
* Cataloges and Shopping have the highest average rating in paid apps.

However,

We see that books in paid apps have high mean rating, however very less in 
free apps. The same in Catalogs.


Q: Which category has the highgest total number of ratings?

```{r echo=FALSE, Bivariate_Plots5, message=FALSE, warning=FALSE, fig.width=12, fig.height=12}
# A histogram showing total number of user ratings and app category.
ggplot(aes(x=prime_genre,y=rating_count_tot,fill=prime_genre), data = df) +
  geom_bar(stat = 'summary', fun.y = mean) + coord_flip() +
  labs (x="Categories", y="Average of total number of user rating",
title="Does mean of total number of user ratings depend on app category?") + 
  facet_wrap(~is_free) + scale_color_manual(values=c("#999999", "#E69F00", "#56B4E9")) + theme_minimal()

```

WoW! The magic begins to appear!
Users don't give feedback or rating to paid apps. However they do in free apps!


Let's check if price may affect on app category?

```{r echo=FALSE, Bivariate_Plots6, message=FALSE, warning=FALSE,fig.width=12, fig.height=12}
# A histogram plot answering  question of 
# Which category is more expensive than other apps?
ggplot(aes(x=prime_genre,y=price,fill=prime_genre), data = df) +
  geom_histogram(stat = 'summary', fun.y = mean) + coord_flip() + 
  labs (x="Categories", y="Average Price",
        title="Which category is more expensive than other apps?") + 
  theme_minimal()
```

Someting strange, Medical category are the most expensive ones! 
Yes the category field can affect on the price of the app!!
Shopping apps have very less price than other apps

```{r echo=FALSE, Bivariate_Plots7, message=FALSE, warning=FALSE}
# line plot showing the relation between current version rating 
# and average user rating for an app.
ggplot(aes(x=user_rating_ver,y=user_rating), data = df) +
  geom_line(stat = 'summary', fun.y = median) + geom_smooth() + 
  labs (x="User rating of recent version", y="Total average user rating",
        title="Does the current version is always have more rating than 
        the total overall rating?")
```

Correlation between current version user rating and total 
overall user rating is 0.7 which leads to a strong positive 
correlation between them as the above plot shows.

```{r echo=FALSE, Bivariate_Plots8, message=FALSE, warning=FALSE,fig.width=8, fig.height=8}
# The relation between app size and price or user_rating
ggplot(aes(x=price,y=size_mb), data = subset(df,df$price<25&&df$price>0)) +
  geom_boxplot(alpha=1/10) + scale_x_continuous(limits = c(0,10),
                                              breaks = seq(0,10,2)) + 
  scale_y_continuous(limits = c(0,1000), breaks = seq(0,1000,100)) + 
  facet_wrap(~is_game) + labs (x="Price", y="Size (MB)",
          title="The relation between app size and price or user_rating")
```

* Most of paid games are less than $5.
* For Games, price decreases when size of the app getting lower, 
however in general apps there are no clear relation between them.

```{r echo=FALSE, Bivariate_Plots9, message=FALSE, warning=FALSE,fig.width=12, fig.height=12}
# A histogram showung the summary of size_mb (MEAN) and user rating.
ggplot(aes(x=user_rating,y=size_mb), data = subset(df,df$user_rating>0)) +
  geom_histogram(stat = 'summary', fun.y = mean) +
  scale_x_continuous(breaks = seq(1.5,4.5,.5),limits = c(1,5))
```

yes! When app size increases, average user rating for the app increases as well.

# Bivariate Analysis

### What are the relationships I observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

* When we look to the relation between price and the total user ratings for an 
app, we find that users rarely give high rating to expensive apps specially
if it's a game. Most af apps ratings are for the apps that are less than USD 50.

* Most of the apps in this dataset have prices less tan $5. 
Let's have a closer look to prices based on content ratings in further analysis.

* We see that books in paid apps have high mean rating, 
However very less in free apps. The opposite in Catalogs. 
That indicates that paid books apps worth the money paid. 
They have good rating in paid apps rather than free ones.

* Medical app category is the most expensive ones! 
Yes the category field can affect on the price of the app!!
* On the other side shopping and finance apps have less price than other apps.

* Does the average user ratings may depend on the app content? 
NO! All content based apps are nearly have 3 in user rating. 
There are some slight diffferences  but not that much to decide that the content
rating may affect on the user  ratings!


### Some observation in the relationships between the other features \
(not the main feature(s) of interest):

* Number of supported languages not strongly affect on the average user rating. 
The smooth line seems to be at the same level of usee rating (4).

* It seems that the total rating always increases with new versions! 

* What is the relation between app size and price or user_rating? 
App size clearly does not affect on the price of the app. 
Games are more stable in price even if the app size being increased. 

* Is the the same regarding the relation between user rating and app size? NO! 
Average user rating increases while app size increases! 
It seems users give good rating for large sized apps!

### What was the strongest relationship you found?
Relation between user rating for the current version and total average of 
user rating was the strongest relation among what we have found so far.

# Multivariate Plots Section

```{r echo=FALSE, Multivariate_Plots, message=FALSE, warning=FALSE,fig.width=8, fig.height=8}
# Showing the relation between total count ratings, 
  # price and user rating
ggplot(aes(x=log10(rating_count_tot),y=price), data = subset(df,df$price<20&df$rating_count_tot),bandwidth= 1) +
  geom_boxplot(aes(color = factor(user_rating))) + 
  labs (y="Price", x="Log10 transformed total count of ratings ",title="The relation between total count ratings, price and user rating")

print("Summary of total number of ratings /n")
summary(log10(df$rating_count_tot))
```

* We can conclude from the plot above that most of good rated apps are cheap and
they also have been rated with large number of users.

```{r echo=FALSE, Multivariate_Plots2, message=FALSE, warning=FALSE,fig.width=12, fig.height=12}
# Showing two plots for size and total number of ratings and price 
# colored by user rating
q1<- ggplot(aes(x=size_mb,y=price), data = subset(df,df$price<25)) +
  geom_point(aes(color = factor(user_rating))) + coord_flip() + labs (y="Price", x="Size (MB)" )

q2<- ggplot(aes(x=log10(rating_count_tot),y=price), 
          data = subset(df,df$price<25))+geom_point(aes(color = factor(user_rating))) + 
          labs (y="Price", x="Transformed log10 of total count of ratings " ) + coord_flip()

grid.arrange(q1,q2,ncol=2)
```

* We see that prices under $10 their sizes vary, 
their user rating stay good while sizes increase. 
* Overall, average user rating remain good though the number 
of total ratings increase.

```{r echo=FALSE, Multivariate_Plots3, message=FALSE, warning=FALSE,fig.width=12, fig.height=12}
# Relation between price and different category colored by user rating to define
# which category has more price in comparing of user rating.
ggplot(aes(x=prime_genre,y=price), data = df) +
  geom_line(aes(color = factor(user_rating))) + 
  theme(axis.text.x=element_text(angle=90,hjust=1)) +
  labs(y="Price",x="Categories")
```

Utilities are expensive apps which have low user ratings.

```{r echo=FALSE, Multivariate_Plots5, message=FALSE, warning=FALSE,fig.width=12, fig.height=12}
# The relation and correlation between user rating and total number of ratings.
ggplot(aes(x=prime_genre,y=log10(rating_count_tot)), data = df) +
  geom_line(aes(color = factor(user_rating))) + 
  labs(y="Total number of ratings",x="Categories") + 
  theme(axis.text.x=element_text(angle=90,hjust=1))

print ("Correlation between user rating and total number of ratings")
cor(df$rating_count_tot,df$user_rating)
```

* We see from the plot above that total rating count have a small effect on 
the user rating for different categories.

```{r echo=FALSE, warning=FALSE, Multivariate_Plots4, message=FALSE,fig.width=12, fig.height=12}
ggplot(aes(x=rating_count_tot,y=user_rating), data = df) +
  geom_point(aes(color = is_free)) + coord_flip()
```

# Multivariate Analysis

Low prices apps are not good enough in general, the user rating increases in  
most of category when the price increases; specially in Navigation, Education 
and References apps. 

* By plotting price and size of the app per MB show that apps have nearly price.
i.e: $5, $10 or $15.
* Low prices of Navigation apps are not good. 
Moderate expensive music apps have good rating though.


* Users love to give good rating when they notice the app already has a good
rating. That's why while total number of rating increases, the average of 
user_rating increases too.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One, message=FALSE, warning=FALSE}
# Plotting final summary
# Prices based on categories
ggplot(aes(x=prime_genre,y=price), data = df) + 
  geom_histogram(stat = 'summary', fun.y = mean) + coord_flip() + 
  labs (x="Category", y="Price",title="Prices based on categories")

# User rating based on categories
ggplot(aes(x=prime_genre,y=user_rating), data = df) +
  geom_histogram(stat = 'summary', fun.y = mean) + coord_flip() +
  labs (x="Category", y="User Rating",title="User rating based on categories") +
  facet_wrap(~is_free) 
```

### Description One
* Medical app category is the most expensive ones, 
may be because they provide valuable information?
* The category field can affect on the price of the app!! 
Shopping and finance apps have less price than other apps. Make sense because 
they provide services and users already pay to use the service.

### Plot Two
```{r echo=FALSE, Plot_Two, message=FALSE, warning=FALSE}
# line plot showing avg User rating vs Rating for recent version
ggplot(aes(x=user_rating_ver,y=user_rating), data = df) + 
  geom_line(stat = 'summary', fun.y = median) + geom_smooth() +
  labs (x="User Rating for last version", y="Avg User Rating",
        title="Avg User rating vs Rating for recent version") 
```

### Description Two
Newer versions of most of the apps have better rating than the median rating. 
Developers always try to publish a better app that worth a better rating always. 
Correlation is 0.7 positive between twose two variables.

### Plot Three
```{r echo=FALSE, Plot_Three, message=FALSE, warning=FALSE,fig.width=12, fig.height=12}
# points plot showing the relation between app price based on size
ggplot(aes(x=size_mb,y=price), data = subset(df,df$user_rating>0)) +
  geom_point(aes(color = factor(user_rating))) + coord_flip() + 
  ylim(limits = c(0,10)) +
  labs (x="Size (MB)", y="Price",title="App price based on size") 
```

### Description Three

* What is the relation between user rating, price and app size? 
Average user rating increases with good reviews while app size increases! 
It seems users give good rating for large sized apps!

* App size clearly does not affect on the price of the app. 
Games are more  stable in price even if the app size being increased. 

------

# Reflection
Our dataset has more than 7000 Apple iOS mobile application details. 
The data was extracted from the iTunes Search API at the Apple Inc website.
The main goal for investigating this data set is to predict if the app details    
(i.e: price, content rating and size) affect on the average user rating for it?

* Categories based on price?
Medical category are the most expensive ones, the category field can affect on 
the price of the app

* Categories based on average user rating?
1. Productivity and Music have the highest average rating in free apps.
2. Cataloges and Shopping have the highest average rating in paid apps.
3. Books in paid apps have high mean rating, however very less in 
free apps. The same in Catalogs. Books paid app worth the money.

* Top 3 apps based on total number of user ratings:
1. Facebook (General App)
2. Instagram (General App)
3. Clash of Clans (Game)

* Top 3 apps based on average user_rating:
1. Domino's Pizza USA (General App)
2. Flashlight (General App)
3. Head Soccer (Game)
4. Plants vs. Zombies (Game)

> **For Future work**: Would it possible to predict success of an app by creating a model? 
We may suppose that app which has more than 4 user_rating is succeful. 
We may also think for a better approach to calculate the ration of a user rating based on total number of ratings count.

```{r echo=FALSE, warning=FALSE, LM, message=FALSE, include=FALSE}
# Creating a linear model to predict app price or user rating for an app.
m1 <- lm(user_rating ~ price , data = df)
m2 <- update(m1, ~ . + price)
m3 <- update(m2, ~ . + prime_genre)
m4 <- update(m3, ~ . + cont_rating)
m5 <- update(m4, ~ . + size_mb)
m6 <- update(m5, ~ . + lang.num)

mtable(m1,m2,m3,m4,m5,m6,sdigits = 3)

xx <- data.frame(prime_genre="Social Networking",
                 user_rating=4,cont_rating="4+",size_mb=100,lang.num=2)
yy <- predict (m6, newdata = xx, interval="prediction",level = .90)
exp(yy)
```
